Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-03-12 paper/paper.tex
restructure

\subsection{Sampling extremely large configuration space}
When the graph size goes large, the configurations might be impossible to fit into any type of storage.
Then instead of using the set algebra, one can use a binary sum-product tree as a compact representation of the configurations.
We use a quadruple $(type, data, left, right)$ to represent a node in the tree, where $type$ is one of
$\{\texttt{LEAF}, \texttt{ZERO}, \texttt{SUM}, \texttt{PROD}\}$, $data$ is a bit string as the content in a \texttt{LEAF} node,
$left$ and $right$ are left and right operands of \texttt{SUM} and \texttt{PROD} nodes.

@@

Figure~\ref{fig:benchmark}(a) shows the time and space complexity of tensor network contraction for different graph sizes, where the space complexity is the same as the treewidth of the problem graph.
The contraction order is obtained using the local search algorithm in Ref.~\cite{Kalachev2021}.
In practice, slicing technique is used for graphs with treewidth greater than 27 to fit the computation into a 16GB memory.
One can see that all the computing times in figure~\ref{fig:benchmark}(b), (c) and (d) have a strong correlation with the treewidth.
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions to evaluate tensor contractions, hence 
are much faster comparing to non-BLAS element types in the same category.
GPU computes much faster than CPU in all cases when the problem scale is large enough such that the actual computing time is comparable or larger than the launching overhead of CUDA kernels.
Most algebras can be computed on a GPU, except those requiring dynamic sized structures, i.e.\ \texttt{PN} and \texttt{SN}.
In figure~\ref{fig:benchmark}(c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
but it may suffer from round-off errors. The finite field (GF$(p)$) approach is the only method that does not have round-off errors and can be run on a GPU.
In figure~\ref{fig:benchmark}(d), one can see the technique to bound the enumeration space improves the performance for more than one order of magnitude in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a device with 32GB main memory.

@@

\section{Additional contraction overhead using the standard tensor network notations}\label{sec:tensorbad}
As we have mentioned in the main text,
a standard tensor network notation is equivalent to the generalized tensor network by introducing $\delta$ tensors,
where a $\delta$ tensor of rank $d$ is defined as
\begin{equation}
    \delta_{i_1, i_2,\ldots,i_d} = \begin{cases}
        1, & i_1=i_2=\ldots =i_d,\\
        0, & \text{otherwise}.
    \end{cases}
\end{equation}

>>>>

\subsection{Sampling extremely large configuration space}
When the graph size goes large, the configurations might be impossible to fit into any type of storage.
Then instead of using the set algebra, one can use a binary sum-product tree as a compact representation of the configurations.
We use a quadruple $(type, data, left, right)$ to represent a node in the tree, where $type$ is one of
\texttt{LEAF}, \texttt{ZERO}, \texttt{SUM} and \texttt{PROD}, $data$ is a bit string as the content in a \texttt{LEAF} node,
$left$ and $right$ are left and right operands of \texttt{SUM} and \texttt{PROD} nodes.

@@

Figure~\ref{fig:benchmark}(a) shows the time and space complexity of tensor network contraction for different graph sizes, where the space complexity is the same as the treewidth of the problem graph.
The contraction order is obtained using the local search algorithm in Ref.~\cite{Kalachev2021}.
In practice, slicing technique is used for graphs with treewidth greater than 27 to fit the computation into a 16GB memory.
One can see that all the computing times in figure~\ref{fig:benchmark}(b), (c) and (d) have a strong correlation with the treewidth.
Among these benchmarks, computational tasks with data types \texttt{T (CPU)},
\texttt{$\mathbbm{R}$ (CPU)}, \texttt{$\mathbbm{R}$ (GPU)}, \texttt{$\mathbbm{C}$ (CPU)}, \texttt{$\mathbbm{C}$ (GPU)} and \texttt{T+bounding (CPU)} can utilize fast BLAS functions to evaluate tensor contractions, hence 
are much faster comparing to non-BLAS element types in the same category.
GPU computes much faster than CPU in all cases when the problem scale is large enough such that the actual computing time is comparable or larger than the launching overhead of CUDA kernels.
Most algebras can be computed on a GPU, except those requiring dynamic sized structures, i.e.\ \texttt{PN}, \texttt{SN}, \texttt{T100} and \texttt{SPT}.
In figure~\ref{fig:benchmark}(c), one can see the Fourier transformation based method is the fastest in computing the independence polynomial,
but it may suffer from round-off errors. The finite field (GF$(p)$) approach is the only method that does not have round-off errors and can be run on a GPU.
In figure~\ref{fig:benchmark}(d), one can see the technique to bound the enumeration space improves the performance for more than one order of magnitude in enumerating the MISs.
Bounding can also reduce the memory usage significantly, without which the largest computable graph size is only $\sim150$ on a device with 32GB main memory.

@@

\section{An example of contraction complexity increase due to introducing $\delta$ tensors}\label{sec:tensorbad}
As we have mentioned in the main text,
a standard tensor network notation is equivalent to the generalized tensor network by introducing $\delta$ tensors,
where a $\delta$ tensor of rank $d$ is defined as
\begin{equation}
    \delta_{i_1, i_2,\ldots,i_d} = \begin{cases}
        1, & i_1=i_2=\ldots =i_d,\\
        0, & \text{otherwise}.
    \end{cases}
\end{equation}
